04-Data-Analysis

### 2021-04-20 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd
Push changes to nb04
Update to include auto passengers

COVIDClose is an indicator variable for whether this bank closed during the pandemic. Convert to factor:
```{r}
foodbank_locations <- foodbank_locations %>%
  mutate(STATUS = replace_na(STATUS, "Open"),
         COVIDClose = if_else(STATUS == "Cancelled", true = "Yes", false = "No")) %>%

@@

Filter modes:
```{r}
travel_modes <- travel_modes %>%
  dplyr::filter(Mode == "Auto driver" |
                  Mode == "Transit (Excluding GO rail)" |
                  Mode == "Walk")
```

@@

Pivot wider and rename columns:
```{r}
travel_modes_1 <- travel_modes_1 %>%
  rename(TAZUID = gta06_hhld) %>%
  pivot_wider(names_from = Mode,
              values_from = trips,
              values_fill = 0) %>%
  rename(Driver = `Auto driver`, 
         Transit = `Transit (Excluding GO rail)`)
```

>>>>

`STATUS` corresponds to the food bank's status during the pandemic. Let's simplify it to a yes-no factor for whether the service remained open or closed after COVID began and then filter out inapplicable services:
```{r}
foodbank_locations <- foodbank_locations %>%
  mutate(STATUS = replace_na(STATUS, "Open"),
         COVIDClose = if_else(STATUS == "Cancelled", true = "Yes", false = "No")) %>%

@@

Filter modes:
```{r}
travel_modes <- travel_modes %>%
  dplyr::filter(Mode == "Auto driver" | 
                  Mode == "Auto passenger" |
                  Mode == "Transit (Excluding GO rail)" |
                  Mode == "Walk")
```

@@

Pivot wider and rename columns:
```{r}
travel_modes_1 <- travel_modes_1 %>%
  rename(TAZUID = gta06_hhld) %>%
  pivot_wider(names_from = Mode,
              values_from = trips,
              values_fill = 0) %>%
  mutate(Driver = `Auto driver` + `Auto passenger`) %>%
  rename(Transit = `Transit (Excluding GO rail)`) %>%
  select(-c(`Auto driver`, `Auto passenger`))
```

### 2021-04-20 00-Data-Processing/04-Data-Analysis.Rmd
Push changes to nb04
Update to include auto passengers

## TEMP: REMOVE INAPPLICABLE DESTINATIONS UNTIL NEW TTMs ARE FINISHED

```{r}
ttm_car <- ttm_car %>%
  filter(!between(OBJECTID, left = 59, right = 69))

ttm_transit <- ttm_transit %>%
  filter(!between(OBJECTID, left = 59, right = 69))

ttm_walk <- ttm_walk %>%
  filter(!between(OBJECTID, left = 59, right = 69))
```

@@

Half of all parcels can reach 52 foodbanks or more in 180 min or less.

@@

Half of all parcels can reach 45 foodbanks or more in 180 min or less.

@@

@@

Join proportion of trips to travel time table:
```{r}
ttm <- ttm %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")
```

@@

@@

### CODE TO BE UPDATED
```{r}
# join tripmaking modal proportions
ttm_car <- ttm_car %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")

Get minimum travel times in the pre-covid scenario:

```{r minimum-travel-time-baseline, include=FALSE}
# retrieve minimum travel time during covid to sites that didnt close
min_ttm_car_pre <- ttm_car %>%
  #dplyr::filter(COVIDClose == "No") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            linc_age_0to2 = first(linc_age_0to2),
            linc_age_0to19 = first(linc_age_0to19),
            linc_age_65plus = first(linc_age_0to19),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")

min_ttm_transit_pre <- ttm_transit %>%
  #dplyr::filter(COVIDClose == "No") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            linc_age_0to2 = first(linc_age_0to2),
            linc_age_0to19 = first(linc_age_0to19),
            linc_age_65plus = first(linc_age_0to19),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")

min_ttm_walk_pre <- ttm_walk %>%
  #dplyr::filter(COVIDClose == "No") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            linc_age_0to2 = first(linc_age_0to2),
            linc_age_0to19 = first(linc_age_0to19),
            linc_age_65plus = first(linc_age_0to19),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")
```

Get minimum travel times in the during covid scenario:

>>>>

## Remove entries in travel time matrices for inapplicable destinations

As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
```{r}
ttm_car <- ttm_car %>%
  filter(!between(OBJECTID, left = 1, right = 3))

ttm_transit <- ttm_transit %>%
  filter(!between(OBJECTID, left = 1, right = 3))

ttm_walk <- ttm_walk %>%
  filter(!between(OBJECTID, left = 1, right = 3))
```

@@

Half of all parcels can reach 49 foodbanks or more in 180 min or less.

@@

Half of all parcels can reach 42 foodbanks or more in 180 min or less.

@@

```{r}
# join tripmaking modal proportions
ttm_car <- ttm_car %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")

@@

Join proportion of trips to travel time table:
```{r}
#ttm <- ttm %>%
#  left_join(modes %>%
#              st_drop_geometry() %>%
#              transmute(TAZUID,
#                        p_car,
#                        p_transit,
#                        p_walk),
#            by = "TAZUID")
```

@@

```{r}
access_by_da <- accessibility %>%
  st_drop_geometry() %>%
  # Copy the DA identifier to the accessibility table
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              transmute(UID = ID, DAUID),
            by = "UID") %>%
  # Group by DA and time
  group_by(DAUID, timing) %>%
  # Calculate the accessibility by DA before and during covid
  summarize(accessibility = sum(accessibility),
            timing = first(timing),
            .groups = "drop") %>%
  # Pivot wider to calculate the differences in accessibility from pre-covid to covid
  pivot_wider(names_from = timing, 
              values_from = accessibility) %>%
  # Calculate changes in accessibility
  mutate(acc_d = `During covid` / `Pre-covid` - 1) %>%
  # Join DA geometry
  left_join(data_da_2016 %>%
            transmute(DAUID = GeoUID,
                      geometry),
          by = "DAUID") %>%
  # Convert to simple features
  st_as_sf() 
tmap_mode("view")
tm_shape(access_by_da) + 
  tm_fill("acc_d")
```

@@

### CODE TO BE UPDATED
Get minimum travel times in the pre-covid scenario:

```{r minimum-travel-time-baseline, include=FALSE}
# retrieve minimum travel time during covid to sites that didnt close
min_ttm_car_baseline <- ttm_car %>%
  #dplyr::filter(COVIDClose == "No") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            linc_age_0to2 = first(linc_age_0to2),
            linc_age_0to19 = first(linc_age_0to19),
            linc_age_65plus = first(linc_age_0to19),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop") %>%

  ## Calculate weighted travel time:
  mutate(weighted_travel_time = travel_time * p_car)

min_ttm_transit_baseline <- ttm_transit %>%
  #dplyr::filter(COVIDClose == "No") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            linc_age_0to2 = first(linc_age_0to2),
            linc_age_0to19 = first(linc_age_0to19),
            linc_age_65plus = first(linc_age_0to19),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop") %>%

  ## Calculate weighted travel time:
  mutate(weighted_travel_time = travel_time * p_transit)

min_ttm_walk_baseline <- ttm_walk %>%
  #dplyr::filter(COVIDClose == "No") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            linc_age_0to2 = first(linc_age_0to2),
            linc_age_0to19 = first(linc_age_0to19),
            linc_age_65plus = first(linc_age_0to19),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop") %>%

  ## Calculate weighted travel time:
  mutate(weighted_travel_time = travel_time * p_walk)
```

```{r weighted-travel-time-by-mode-baseline, include=FALSE}
# Baseline analysis (pre-COVID)
## Bind the three tables and label them by mode:
min_ttm_mode_baseline <- rbind(data.frame(min_ttm_car_baseline, Mode = "Car"),
                               data.frame(min_ttm_transit_baseline, Mode = "Transit"),
                               data.frame(min_ttm_walk_baseline, Mode = "Walking"))
```

```{r aggregate-weighted-travel-times-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Aggregate weighted travel times by mode:
min_ttm_baseline <- min_ttm_mode_baseline %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            linc_age_0to2 = first(linc_age_0to2),
            linc_age_0to19 = first(linc_age_0to19),
            linc_age_65plus = first(linc_age_0to19),
            weighted_travel_time = sum(weighted_travel_time)) %>%

  ## Calculate person-hours of travel:
  mutate(weighted_person_hours_0to2 = linc_age_0to2 * weighted_travel_time/60)
```

```{r summarize-results-by-taz-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Summarize by TAZ and join geometry:
min_ttm_baseline_taz <- min_ttm_baseline %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            weighted_person_hours_0to2 = sum(weighted_person_hours_0to2)) %>%
  left_join(modes %>%
              dplyr::select(TAZUID),
            by = "TAZUID") %>%
  st_as_sf()
```

```{r weighted-travel-time-map-baseline}
ttm_map_baseline <- ggplot() +
  geom_sf(data = min_ttm_baseline_taz,
          aes(fill = weighted_travel_time),
          color = NA) +
  scale_fill_distiller(name = "Expected travel time (min)",
                       palette = "OrRd", 
                       direction = 1) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_text(size = 7),
        legend.position = "bottom",
        legend.text = element_text(size = 8))

ttm_map_baseline
```

```{r}
tmap_mode("view")
tm_shape(min_ttm_baseline_taz) +
  tm_fill(col = "weighted_travel_time")

tm_shape(modes) +
  tm_fill(col = "p_walk")
```

```{r person-hours-traveled-map-baseline}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Map weighted person-hours of travel:
pht_map_baseline <- ggplot() +
  geom_sf(data = min_ttm_baseline_taz,
          aes(fill = weighted_person_hours_0to2),
          color = NA) +
  scale_fill_distiller(name = "Person-hours of travel",
                       palette = "OrRd", 
                       direction = 1) +
  #geom_sf(data = urban_types,
  #        aes(color = Type),
  #        fill = NA) +
  #scale_color_manual(name = "",
  #                   values = c("Urban" = "firebrick2",
  #                              "Suburban" = "cornflowerblue",
  #                              "Rural" = "forestgreen"))  +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_text(size = 7),
        legend.position = "bottom",
        legend.text = element_text(size = 8))

pht_map_baseline
```

```{r join-income-information-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Join median household information to table:
min_ttm_baseline <- min_ttm_baseline %>%
  left_join(data_da_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(DAUID = GeoUID,
                               income_quintile),
            by = "DAUID")
```

```{r plot-distribution-travel-time-by-income-baseline}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Distribution of weighted travel time by income:
ttm_income_baseline <- ggplot(data = min_ttm_baseline, 
       aes(x = weighted_travel_time, 
           color= income_quintile)) +
  geom_density(size = 0.5, adjust = 2.5) +
  labs(x = "Travel time (in minutes; weighted by mode)") +
  scale_color_manual(values = c("Top 20%" = "blue",
                                "Second 20%" = "steelblue1",
                                "Third 20%" = "gray40", 
                                "Fourth 20%" = "tomato",
                                "Bottom 20%" = "red")) +
  theme_minimal() +
  theme(text = element_text(size = 7),
        legend.text = element_text(size = 7))

ttm_income_baseline
```

```{r total-weighted-person-hours-by-group-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Total weighted person-hours by income:
person_hours_income_baseline <- min_ttm_baseline %>%
  group_by(income_quintile) %>%
  summarize(total_weighted_person_hours = sum(weighted_person_hours),
            total_population = sum(Population),
            .groups = "drop") %>%
  mutate(ratio = total_weighted_person_hours/total_population) %>%
  rename(group = income_quintile)

## Total weighted person-hours by region:
person_hours_region_baseline <- min_ttm %>%
  group_by(Urban_type) %>%
  summarize(total_weighted_person_hours = sum(weighted_person_hours),
            total_population = sum(Population),
            .groups = "drop") %>%
  drop_na() %>%
  mutate(ratio = total_weighted_person_hours/total_population) %>%
  rename(group = Urban_type)

person_hours_baseline <- rbind(person_hours_income_baseline,
                               person_hours_region_baseline)
```

### Do Again for COVID Scenario
Get minimum travel times in the during covid scenario:
